---
title: "SITS-AMZ Pipeline"
output:
  html_document:
    self_contained: true
    css: styles.css
---

### Introduction

This flowchart shows the pipeline used in the PRODES-AMZ project, developed with the SITS package. The data and results presented here are available in the repository.

### Description

1.  [Creating Linear Spectral Mixture Model (LSMM) for features training and classification.](https://github.com/migualex/sits-prodes/blob/main/scripts/1_SITS_mixture_model.R)
2.  [Segmentation from Linear Spectral Mixture Model (LSMM) for features training and classification.](https://github.com/migualex/sits-prodes/blob/main/scripts/2_SITS_segmentation.R)
3.  [Samples quality assessment, filtering and balancing.](https://github.com/migualex/sits-prodes/blob/main/scripts/3_SITS_samples_quality_assessment.R)
4.  [Train Machine Learning Model.](https://github.com/migualex/sits-prodes/blob/main/scripts/4_SITS_model_training.R)
5.  [Classification of Vector Data Cube.](https://github.com/migualex/sits-prodes/blob/main/scripts/5_SITS_classification.R)
6.  [Stratification of validation samples.](https://github.com/migualex/sits-prodes/blob/main/scripts/6_SITS_validation_samples_stratification.R)
7.  [Validation and accuracy measurement.](https://github.com/migualex/sits-prodes/blob/main/scripts/7_SITS_validation_accuracy_measurement.R)
8.  [Post-Processing.](https://github.com/migualex/sits-prodes/blob/main/scripts/7_SITS_post_processing.R)

```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(DiagrammeR)
library(DiagrammeRsvg)

sits_pipeline <- grViz("
digraph SITS_AMZ_PIPELINE {

  graph [
    layout = dot,
    rankdir = TB,
    splines = ortho,
    fontsize = 18,
    fontname = Helvetica,
    labelloc = t
  ]

  node [
    shape = box,
    style = filled,
    fontname = Helvetica,
    fontsize = 11
  ]

  edge [
    fontname = Helvetica,
    fontsize = 10,
    color = gray30,
    arrowhead = vee
  ]
  
  // -------------------------
  // Legend
  // -------------------------
  # subgraph cluster_legend {
  #   label = <<B>Legend</B>>
  #   style = rounded
  #   color = gray50
  #   fontname = Helvetica
  #   fontsize = 14
  #   
  # node_cube   [fillcolor=\"#E1F5FE\", color=\"#01579B\"]
  # node_func   [fillcolor=\"#FFF9C4\", color=\"#F9A825\"]
  # node_file   [fillcolor=\"#E8F5E9\", color=\"#2E7D32\", style=\"filled,dashed\"]
  # node_logic  [shape=diamond, fillcolor=\"#F3E5F5\", color=\"#7B1FA2\"]
  # }
  
  subgraph cluster_script1 {
    label = <<B>Step 1</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S1_Cube [label=\"Load Sentinel-2 Cube\\nsits_cube()\", fillcolor=\"#E1F5FE\"]
    S1_Endm [label=\"Define Endmembers\\nSoil / Veg / Dark\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S1_LSMM [label=\"Generate LSMM Fractions\\nsits_mixture_model()\", fillcolor=\"#FFF9C4\"]
    S1_Out  [label=\"LSMM Cube\\nSoil / Veg / Dark\", fillcolor=\"#E1F5FE\"]
    S1_Dir  [label=\"Saved LSMM cube\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S1_Cube -> S1_LSMM
    S1_Endm -> S1_LSMM
    S1_LSMM -> S1_Out
    S1_LSMM -> S1_Dir
  }

  subgraph cluster_script2 {
    label = <<B>Step 2</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S2_LoadMM [label=\"Load LSMM cube\\nsits_cube(data_dir)\", fillcolor=\"#E1F5FE\"]
    S2_SNIC   [label=\"Segment LSMM cube\\nsits_segment()\\nseg_fn = sits_snic()\", fillcolor=\"#FFF9C4\"]
    S2_SegVec [label=\"Saved segments\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S2_LoadMM -> S2_SNIC -> S2_SegVec
  }

  subgraph cluster_script3 {
    label = <<B>Step 3</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S3_CubeS2   [label=\"Load Training Cube\\nsits_cube()\", fillcolor=\"#E1F5FE\"]
    S3_LoadMM   [label=\"Load LSMM cube\\nSoil / Veg / Dark\", fillcolor=\"#E1F5FE\"]
    S3_Merge    [label=\"Merge features\\nsits_merge()\", fillcolor=\"#FFF9C4\"]

    S3_Samples  [label=\"Load Training Samples\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S3_GetData  [label=\"Extract time series\\nsits_get_data()\", fillcolor=\"#FFF9C4\"]
    S3_TS       [label=\"Time Series\", fillcolor=\"#E1F5FE\"]
    S3_SaveTS   [label=\"Save original\\nsaveRDS(time_series)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S3_SOM1     [label=\"SOM clustering (1)\\nsits_som_map()\", fillcolor=\"#FFF9C4\"]
    S3_Eval1    [label=\"Evaluate clusters\\nsits_som_evaluate_cluster()\", fillcolor=\"#FFF9C4\"]
    S3_Clean    [label=\"Clean samples\\nsits_som_clean_samples()\", fillcolor=\"#FFF9C4\"]

    S3_SOM2     [label=\"SOM clustering (2)\\nsits_som_map(cleaned)\", fillcolor=\"#FFF9C4\"]
    S3_Eval2    [label=\"Evaluate clusters\\nsits_som_evaluate_cluster(cleaned)\", fillcolor=\"#FFF9C4\"]

    S3_Balance  [label=\"Reduce imbalance\\nsits_reduce_imbalance()\", fillcolor=\"#FFF9C4\"]
    S3_SaveBal  [label=\"Save cleaned & balanced\\nsaveRDS(new_time_series)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S3_SOM3     [label=\"SOM clustering (3)\\nsits_som_map(cleaned + balanced)\", fillcolor=\"#FFF9C4\"]
    S3_Eval3    [label=\"Evaluate clusters\\nsits_som_evaluate_cluster(cleaned + balanced)\", fillcolor=\"#FFF9C4\"]

    S3_CubeS2 -> S3_Merge
    S3_LoadMM -> S3_Merge

    S3_Samples -> S3_GetData
    S3_Merge   -> S3_GetData

    S3_GetData -> S3_TS -> S3_SaveTS

    S3_TS -> S3_SOM1 -> S3_Eval1 -> S3_Clean
    S3_Clean -> S3_SOM2 -> S3_Eval2
    S3_Clean -> S3_Balance -> S3_SaveBal
    S3_SaveBal -> S3_SOM3 -> S3_Eval3
  }

  subgraph cluster_script4 {
    label = <<B>Step 4</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S4_LoadTS [label=\"Load training\\nreadRDS(cleaned & balanced)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S4_Kfold  [label=\"Cross-validation\\nsits_kfold_validate()\", fillcolor=\"#FFF9C4\"]
    S4_Train  [label=\"Train Model\\nsits_train()\", fillcolor=\"#FFF9C4\"]
    S4_Model  [label=\"Trained Model\", fillcolor=\"#E1F5FE\"]
    S4_Save   [label=\"Save model\\nsaveRDS()\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S4_LoadTS -> S4_Kfold -> S4_Train -> S4_Model
    S4_Model -> S4_Save
  }

  subgraph cluster_script5 {
    label = <<B>Step 5</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S5_Cube    [label=\"Load classification cube\\nsits_cube()\", fillcolor=\"#E1F5FE\"]
    S5_LoadMM  [label=\"Load LSMM fractions\\nsits_cube()\", fillcolor=\"#E1F5FE\"]
    S5_Merge   [label=\"Merge features\\nsits_merge()\", fillcolor=\"#FFF9C4\"]
    S5_Seg     [label=\"Load segmented vector cube\\nsits_cube(vector_dir)\", fillcolor=\"#E1F5FE\"]
    S5_Model   [label=\"Load RF model\\nreadRDS()\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S5_Class   [label=\"Classify segments\\nsits_classify()\", fillcolor=\"#FFF9C4\"]
    S5_Label   [label=\"Generate final class map\\nsits_label_classification()\", fillcolor=\"#FFF9C4\"]
    S5_Map     [label=\"Classified vector map\\n(class_*.gpkg)\", fillcolor=\"#E1F5FE\"]
    S5_Unc     [label=\"Entropy uncertainty\\nsits_uncertainty()\", fillcolor=\"#FFF9C4\"]
    S5_UncOut  [label=\"Uncertainty raster\\n(entropy_*.tif)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S5_Cube -> S5_Merge
    S5_LoadMM -> S5_Merge
    S5_Merge -> S5_Seg

    S5_Model -> S5_Class
    S5_Seg -> S5_Class

    S5_Class -> S5_Label -> S5_Map
    S5_Map -> S5_Unc -> S5_UncOut
  }

  subgraph cluster_script6 {
    label = <<B>Step 6</B>>
    style = rounded
    color = gray50
    labeljust = \"l\"
    fontname = Helvetica
    fontsize = 14

    S6_Rast    [label=\"Rasterize classified vectors\\nsits_rasterize_segments()\", fillcolor=\"#FFF9C4\"]
    S6_Cube    [label=\"Load classified raster cube\\nsits_cube(bands=class)\", fillcolor=\"#E1F5FE\"]
    S6_Desi    [label=\"Sampling Design\\nsits_sampling_design()\", fillcolor=\"#FFF9C4\"]
    S6_Strat   [label=\"Stratified random sampling\\nsits_stratified_sampling()\", fillcolor=\"#FFF9C4\"]
    S6_Val     [label=\"Validation samples\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S6_Acc     [label=\"Accuracy assessment\\nsits_accuracy()\", fillcolor=\"#FFF9C4\"]
    S6_CM      [label=\"Confusion matrix\", fillcolor=\"#E1F5FE\"]
    S6_ME      [label=\"Metrics\", fillcolor=\"#E1F5FE\"]

    S6_Rast -> S6_Cube -> S6_Desi
    S6_Desi -> S6_Strat -> S6_Val
    S6_Val -> S6_Acc
    S6_Cube -> S6_Acc
    S6_Acc -> S6_CM
    S6_Acc -> S6_ME
  }

  subgraph cluster_script7 {
    label = <<B>Step 7</B>>
    style = rounded
    color = gray50
    fontname = Helvetica
    fontsize = 14

    S7_Class    [label=\"Load classified vectors\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S7_Filter   [label=\"Filter deforestation classes\", fillcolor=\"#FFF9C4\"]
    S7_Dissolve [label=\"Dissolve + explode polygons\\nst_union() + st_cast()\", fillcolor=\"#FFF9C4\"]
    S7_Prodes   [label=\"Load PRODES mask\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]
    S7_Diff     [label=\"Apply mask difference\\nst_difference()\", fillcolor=\"#FFF9C4\"]
    S7_Area     [label=\"Area threshold\\n≥ 1 ha\", shape=diamond, fillcolor=\"#F3E5F5\"]
    S7_Spatial  [label=\"Spatial rule\\nkeep tree remnants\\n(if not disjoint)\", shape=diamond, fillcolor=\"#F3E5F5\"]
    S7_Final    [label=\"Final post-processed output\\n(gpkg)\", style=\"filled,dashed\", fillcolor=\"#E8F5E9\"]

    S7_Class -> S7_Filter -> S7_Dissolve
    S7_Prodes -> S7_Diff
    S7_Dissolve -> S7_Diff -> S7_Area -> S7_Spatial -> S7_Final
  }

  ROUTE_MM [shape=point, width=0.01, label=\"\", style=invis]
  ROUTE_ML [shape=point, width=0.01, label=\"\", style=invis]

  S1_Out -> ROUTE_MM [arrowhead=none, color=gray35]
  ROUTE_MM -> S2_LoadMM
  ROUTE_MM -> S3_LoadMM

  S3_SaveBal -> S4_LoadTS
  S4_Save -> S5_Model
  
  S5_Map -> ROUTE_ML [arrowhead=none, color=gray35]
  ROUTE_ML -> S6_Rast
  ROUTE_ML -> S7_Class
}
")

sits_pipeline
```
<div class="center-diagram">
```{r echo=FALSE}
svg <- DiagrammeRsvg::export_svg(sits_pipeline)
writeLines(svg, "C:/Users/usuario/Documents/sits-prodes/docs/pipeline_flowchart.svg")
```
</div>